---
title: "saliva_stability"
author: "Dan Weinberger"
date: "4/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(reshape2)
library(readxl)
```

Freeze-thaw Key
Glycerol: 1 - with glycerol
Glycerol: 0 - without glycerol
Temperature: 2  -80ºC
Temperature: 1  -20ºC
Freeze-thaw Cycle: 0-3
Concentration: CFU/mL
Time: Hours


Timecourse Key
Temperature: 1 - 4ºC
Temperature: 2-  Room temperature
Time: Hours
Concentration: CFU/mL


```{r}
a1 <- read_excel('./Data/Spn stability analysis for R.xlsx')
a1$piab <- as.numeric(a1$piab)

a1 <- a1[a1$Temperature %in% c(1,2),]

a1$Serotype <- as.factor(a1$Serotype)
a1$cycle <- as.factor(a1$cycle)
a1$Temperature<- as.factor(a1$Temperature)
a1$glycerol<- as.factor(a1$glycerol)

a1$concentration<- as.factor(a1$concentration)

```

```{r}

ggplot(a1, aes(x=cycle, y=piab, group=Serotype, color=Serotype))+
  geom_point(aes(shape=concentration)) +
  theme_classic() +
  facet_grid(glycerol~Serotype) +
  geom_hline(yintercept = 40, lty=2, col='gray')
```

Base model

```{r}
mod1 <- lm(piab ~ Serotype + Temperature + glycerol +cycle +concentration, data=a1)

summary(mod1)
```
Does the inclusion of glycerol to the saliva samples make a difference to viability of the pneumococcus? Doesn't appear to

```{r}
mod2 <- lm(piab ~ Serotype + Temperature + concentration+ glycerol +cycle + glycerol*cycle , data=a1)

summary(mod2)

# mod2a <- lm(piab ~ Serotype + Temperature + concentration+ glycerol +cycle + concentration*cycle , data=a1)
# 
# summary(mod2a)

# mod3 <- lm(piab ~ Serotype + Temperature + concentration+ glycerol +cycle + glycerol*concentration*cycle , data=a1)
# 
# summary(mod3)
```

Does temp affect stability 
```{r}
mod4 <- lm(piab ~ Serotype + Temperature + concentration+ glycerol +cycle + Temperature*cycle , data=a1)

summary(mod4)
```
Does serotype affect survival
```{r}
mod4 <- lm(piab ~ Serotype + Temperature + concentration+ glycerol +cycle + Serotype*cycle , data=a1)

summary(mod4)
```



```{r}
b1 <- read_excel('./Data/Spn stability analysis for R.xlsx', sheet='Timecourse')

b1$piaB <- as.numeric(b1$piaB)

b1 <- b1[b1$Temperature %in% c(1,2),]

b1$Serotype <- as.factor(b1$Serotype)
b1$Temperature<- as.factor(b1$Temperature)
b1$Time <- as.factor(b1$Time)
b1$concentration<- as.factor(b1$concentration)
```


```{r}

ggplot(b1, aes(x=Time, y=piaB, group=Serotype, color=Serotype))+
  geom_point(aes(shape=Temperature)) +
  theme_classic() +
  facet_grid(concentration~Serotype) +
  geom_hline(yintercept = 40, lty=2, col='gray')
```
## Base model for time course; 
shows on average loss of 1 Ct at 24 hours, 3.1 Ct at 48 and 72 hours
```{r}
tmod1 <- lm(piaB ~ Temperature + Serotype + concentration + Time, data=b1)

summary(tmod1)
```
11A is the reference serotype here. At 48 hours, the CT loss for 19F was worse than 11A. Nothing else sticks out
```{r}
tmod2 <- lm(piaB ~ Temperature + Serotype + concentration + Time + Time*Serotype, data=b1)

summary(tmod2)
```
The loss in Ct was not worse for lower starting concentration 
```{r}
tmod3 <- lm(piaB ~ Temperature + Serotype + concentration + Time + Time*concentration, data=b1)

summary(tmod3)
```

The loss in Ct was not worse by Temprature.
```{r}
tmod4 <- lm(piaB ~ Temperature + Serotype + concentration + Time + Time*Temperature, data=b1)

summary(tmod4)
```






